---
output: 
  github_document:
    pandoc_args: --webtex=http://chart.apis.google.com/chart?cht=tx&chl=
bibliography: dependencies.bib   
---

<!--  use the --webtex argument in the YAML to render equations -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning = FALSE, 
  results = "asis",
  encoding ='UTF-8'
)
```


<!-- badges: start -->
<!-- [![License: GPL v3](https://img.shields.io/badge/License-GPL%20v3-blue.svg)](http://www.gnu.org/licenses/gpl-3.0) -->
<!-- [![DOI](https://zenodo.org/badge/240023228.svg)](https://zenodo.org/badge/latestdoi/240023228) -->
<!-- [![Travis build status](https://travis-ci.org/MartinSchobben/point.svg?branch=master)](https://travis-ci.org/MartinSchobben/point) -->
<!-- badges: end -->


# Reproducing the paper

``` r
# Install bookdown
install.packages("bookdown")
```

Knit the paper

``` r
# Knit
rmarkdown::render_site(input = "paper/preprint", encoding = 'UTF-8')
```

```{r pkgs echo=FALSE, message=FALSE, eval=FALSE}


```


## Credits

The construction of the R [@rversion] package *point* and associated documentation was aided by the packages; *devtools* [@devtools], *roxygen2* [@roxygen2], *testthat* [@testthat], *knitr* [@knitr1 ; @knitr2], *rmarkdown* [@rmarkdown1; @rmarkdown2], and the superb guidance in the book: *R packages: organize, test, document, and share your code*, by @Wickham2015. In addition, this package relies on a set of external packages from the tidyverse universe, including: *dplyr* [@dplyr], *tidyr* [@tidyr], *tibble* [@tibble], *stringr* [@stringr], *readr* [@readr], *magrittr* [@magrittr], *ggplot2* [@ggplot2], *rlang* [@rlang], and *purrr* [@purrr] for internal functioning as well as specialised statistics; *polyaAeppli* [@polyaAeppli].


